---
title: "DATA 621 - Homework #1"
author: "Cheryl Bowersox, Christopher Martin, Robert Sellers, Edwige Talla Badjio"
date: "June 19, 2016"
output:
  pdf_document:
    fig_caption: yes
    highlight: pygments
    latex_engine: xelatex
    number_sections: yes
    toc: yes
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
csl: mee.csl
title2: Multiple Linear Regression
---

```{r message=FALSE, echo=FALSE}

#install packages
#this will not be shown

suppressWarnings(library(faraway))
suppressWarnings(library(ggplot2))
suppressWarnings(library(lubridate))
library(dplyr)
library(tidyr)
library(RCurl)
library(rjson)
library(gridExtra)
library(lattice)

#load data
trainingdata <- read.csv("https://raw.githubusercontent.com/RobertSellers/dataWarehouse/master/moneyball-training-data.csv", header=TRUE)
#removing Index
trainingdata<-trainingdata%>%select(-INDEX)
evaluationData <-read.csv("https://raw.githubusercontent.com/RobertSellers/dataWarehouse/master/moneyball-evaluation-data.csv", header=TRUE)
```

#OBJECTIVE

To build an optimal multiple linear regression model for annual number of wins per team based on predictors. The following is a procedure for the selection of three different regression models and the selection of a preferred method.  The data source is of data statistics.



***


#DATA EXPLORATION

```{r echo= FALSE, message=FALSE}
n<-nrow(trainingdata)
meanData<-mean(trainingdata$TARGET_WINS)
sdData<-sd(trainingdata$TARGET_WINS)

```

The data has `r nrow(trainingdata)` rows, with each record representing a professional baseball team from the years 1871 to 2006 inclusive. Each record
has the performance of the team for the given year, with all of the statistics adjusted to match the performance of a 162 game season.

The response variable is the number of wins, or TARGET_WINS, which has an approximate mean of `r round(meanData,1)` and standard deviation of `r round(sdData,1)`. It follows a relatively normal, uniform distribution and we will assume there are no outliers.


```{r echo=FALSE}
old.par <- par(mfrow=c(1, 2))
plot(trainingdata$TARGET_WINS, main="Total Wins Distribution",ylab="Total Wins")
hist(trainingdata$TARGET_WINS,main="Total Wins Histogram",xlab="Total Wins")
par(old.par)

#Cheryl Plots
#Density Plots - looking at distribution of interesting variables

g1 <- ggplot(data = trainingdata, aes(x = TARGET_WINS) )+ 
  geom_density(alpha = .2, fill = "003333")+
  ggtitle("Target Wins")


g2 <- ggplot(data = trainingdata, aes(x = TEAM_BATTING_H) )+ 
  geom_density(alpha = .2, fill = "#CC6666")+
  ggtitle("Base hits by batters")


g3 <- ggplot(data = trainingdata, aes(x = TEAM_BATTING_BB) )+ 
  geom_density(alpha = .2, fill = "#66CC00")+
  ggtitle("Walks by Batters")

g4 <- ggplot(data = trainingdata, aes(x = TEAM_PITCHING_HR) )+ 
  geom_density(alpha = .2, fill = "#9933CC")+
  ggtitle("Home Runs Allowed")

grid.arrange(g1,g2,g3,g4, ncol=2, top ="Density Plots")

#Box Plots

g1<- ggplot(data = trainingdata, aes(x=1, y = TARGET_WINS)) + geom_boxplot(fill = "white", colour = "#3399CC", 
                                                outlier.colour = "red", outlier.shape = 1)+
  xlab("") + ylab("wins") +
  ggtitle("Wins")

g2<- ggplot(data = trainingdata, aes(x=2, y = TEAM_BATTING_H)) + geom_boxplot(fill = "white", colour = "#CC6666", 
                                                         outlier.colour = "blue", outlier.shape = 1)+
  xlab("") + ylab("Batting H") +
  ggtitle("Base hits by batters")

g3<- ggplot(data = trainingdata, aes(x=2, y = TEAM_BATTING_BB)) + geom_boxplot(fill = "white", colour = "#66CC00", 
                                                            outlier.colour = "red", outlier.shape = 1)+
  xlab("") + ylab("Walks") +
  ggtitle("Walks by Batters")

g4<- ggplot(data = trainingdata, aes(x=2, y = TEAM_PITCHING_HR)) + geom_boxplot(fill = "white", colour = "#9933CC", 
                                                           outlier.colour = "red", outlier.shape = 1)+
  xlab("") + ylab("HR allowed") +
  ggtitle("Pitching - HR Allowed")


grid.arrange(g1,g2,g3,g4, ncol=2, top ="Boxplots")

```

The explanatory/predictor variables are the remaining variables that may have a positive or negative impact on the number of wins. 

#DATA PREPARATION 

There are several NAs which appear to be values that should have a 0 value (for example, it is perfectly valid for there to be no field double plays in a given match, or no batters hit by pitch in a given match):

```{r}
trainingdata[is.na(trainingdata)] <- 0
#this instead?  Other than this, we just run the models to not include NA?
#trainingdata$TEAM_BASERUN_CS[is.na(trainingdata)] <- 0
```

For columns with values for HBP (hit by pitch) and DP (double plays), the minimum values are >0 and there are a large number of missing values (10% of DP's and 92% of HBP's). While many of these values could very well be missing, it is highly likely that there are 0 values. To simplify this process, a value of 0 has been given to these missing values, while other columns containing missing values (CS, SB, SO) are removed from once the model is run.

```{r}
#Perhaps only changing those without 0 values to 0 values, then removing the rest when the model runs
trainingdata$TEAM_BATTING_HBP[is.na(trainingdata$TEAM_BATTING_HBP)] <- 0
trainingdata$TEAM_FIELDING_DP[is.na(trainingdata$TEAM_FIELDING_DP)] <- 0

#in TEAM_BASERUN_SB there are only 2 0 values (2 matches with no stolen bases), with the next lowest being 14 stolen bases, and 131 missing values
trainingdata$TEAM_BASERUN_SB[is.na(trainingdata$TEAM_BASERUN_SB)] <- 0

#in TEAM_BASERUN_CS there is only 1 0 values (1 match with a caught stealing), with the next lowest being 7 caught, and 772 missing values
trainingdata$TEAM_BASERUN_CS [is.na(trainingdata$TEAM_BASERUN_CS)] <- 0

#in TEAM_BASERUN_SB there is only 1 0 values (1 match with a caught stealing), with the next lowest being 7 caught, and 772 missing values
trainingdata$TEAM_BASERUN_SB[is.na(trainingdata$TEAM_BASERUN_SB)] <- 0

#in TEAM_PITCHING_SO and BATTING_SO there are 20 0 values (20 matches with strikeouts), with the next lowest being 181/66 strikeouts, and 102 missing values -- likely missing values (not 0)
```

Additionally, since (in baseball) the attacking team is the team whose turn it is at bat it might be interesting to see their success rates for those occurrences where a successful hit-at-bat occurs:

```{r}
'add column for total non-homerun hits (all base hits)'
trainingdata$TEAM_BATTING_HITSBASE <- trainingdata$TEAM_BATTING_H + trainingdata$TEAM_BATTING_2B + trainingdata$TEAM_BATTING_3B

'add column for all base situations'
trainingdata$TEAM_BATTING_ALLBASE <- trainingdata$TEAM_BATTING_H + trainingdata$TEAM_BATTING_2B + trainingdata$TEAM_BATTING_3B + trainingdata$TEAM_BATTING_BB + trainingdata$TEAM_BATTING_HBP

#TEAM_BATTING_H was reduced to become TEAM_BATTING_1B, as the combination of base hits simplifies hits unnecessarily. These data points are best split, as they exist for doubles, triples, and home runs.

trainingdata$TEAM_BATTING_1B<-trainingdata$TEAM_BATTING_H- trainingdata$TEAM_BATTING_2B- trainingdata$TEAM_BATTING_3B- trainingdata$TEAM_BATTING_HR

#TOTAL_BASES'
trainingdata$TOTAL_BASES<-(trainingdata$TEAM_BATTING_2B*2)+(trainingdata$TEAM_BATTING_3B*3)+(trainingdata$TEAM_BATTING_HR*4)+ trainingdata$TEAM_BASERUN_SB+trainingdata$TEAM_BATTING_BB+trainingdata$TEAM_BATTING_1B

#add column for all positive batting situations (base + homerun)'
trainingdata$TEAM_BATTING_ALLPOS <- trainingdata$TEAM_BATTING_H + trainingdata$TEAM_BATTING_BB + trainingdata$TEAM_BATTING_HBP

#add column for all positive attacking situations (base + homerun + steals)'
trainingdata$TEAM_BATTING_ALLPOSATTCK <- trainingdata$TEAM_BATTING_H + trainingdata$TEAM_BATTING_2B + trainingdata$TEAM_BATTING_3B + trainingdata$TEAM_BATTING_HR + trainingdata$TEAM_BATTING_BB + trainingdata$TEAM_BATTING_HBP + trainingdata$TEAM_BASERUN_SB
```

There is also another team playing: the defending team. The defending team is mostly effective with the pitcher, so some added variables for pitching could be interesting:

```{r}
'add column for negative pitches (bases)'
trainingdata$TEAM_PITCHING_BASESGIVEN <- trainingdata$TEAM_PITCHING_H + trainingdata$TEAM_PITCHING_BB

'add column for negative pitching (bases + HRs given)'
trainingdata$TEAM_PITCHING_BASESGIVEN <- trainingdata$TEAM_PITCHING_H + trainingdata$TEAM_PITCHING_BB + trainingdata$TEAM_PITCHING_HR

'add column for success in offence vs failures in pitching'
trainingdata$TEAM_POSRATIO <- trainingdata$TEAM_BATTING_ALLPOSATTCK / trainingdata$TEAM_PITCHING_BASESGIVEN
```

Stolen Bases and caught stolen would be better combined into "Stolen Base Percentage".

```{r}
trainingdata$SBPERCENT<-trainingdata$TEAM_BASERUN_SB/
(trainingdata$TEAM_BASERUN_SB+trainingdata$TEAM_BASERUN_CS)
trainingdata$TEAM_BASERUN_SB<-NULL
trainingdata$TEAM_BASERUN_CS<-NULL
```




***

#BUILD MODELS

```{r echo=FALSE, message=FALSE, results='hide'}
#Multiple linear regression using stepwise methodology
trainingdata<-na.omit(trainingdata)
all<-lm(TARGET_WINS~.,trainingdata)
all<-update(all,~.-INDEX)
null=lm(TARGET_WINS~1, data=trainingdata)
stepResults<-step(null,scope=list(lower=null,upper=all),direction="forward")
stepResults
```

Current optimal results from stepwise method

```{r}
stepResults
```






***

#SELECT MODELS 




***

#APPENDIX